---
title: "Crawler Validation Matrix"
author: Carlos Paradis
output: html_notebook
---

# Introduction

This is a very quickly done notebook just to create a table alike to http://seclists.org/bugtraq to verify the crawler download all e-mails. 

```{r}
library(data.table)
```

On MEGA, the .raw corpus is inside "New Crawler Full Disclosure" and "New Crawler Bugtraq".

To test the **Crawler** works with this Notebook, **unzip** the **<year>.raw.zip** files from all years into a `corpus_folder`. This will create a folder for every year, and inside of it a folder for every month, and inside of it, a list of all e-mails as separate .txt files, and one extra index file necessary to extract the metadata in combination with the .txt files.

See the table in the end and compare it to the one available in seclists. All numbers must match. 

```{r}


corpus_path <- "~/Desktop/corpus_folder" #you should unzip all different <year>.raw.zip into this folder.
year_folders <- list.files (corpus_path)
```

```{r}
counts <- array()
month_paths <- array()

for(i in 1:length(year_folders)){
  year_folder_path <- paste0("~/Desktop/corpus_folder/",year_folders[i])
  month_path <- list.files(year_folder_path)
  month_paths <- c(month_paths,month_path)
  
  for (j in 1:length(month_path)){
    file_paths <- paste0(year_folder_path,"/",month_path[j])
    count <- length(list.files(file_paths)) - 1 #1 file is always the index in the crawler
    counts <- c(counts,count)
  }
}
month_paths <- month_paths[2:length(month_paths)]
counts <- counts[2:length(counts)]
month_paths
```
```{r}
counts
```


```{r}
year_range <- 2005:2016

validation.matrix <- t(matrix(data=counts,nrow=length(year_range),ncol=12))
colnames(validation.matrix) <- as.character(c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
rownames(validation.matrix) <- year_range


validation.matrix <- validation.matrix[ order(-as.numeric(row.names(validation.matrix))), ]
validation.matrix
```

